Geometric variational problems

### 2020-11-10 NSF2020/Description/background.tex
working on page 1

\noindent
{\bf Collaborative Research: Mathematical modeling and simulations of
self-assembling amphiphilic particles in a solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\it Rolf Ryham (lead PI, Fordham University)}, {\it Bryan Quaife (PI,
FSU), and Y.-N. Young (PI, NJIT)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to investigate the dynamic self-assembly of amphiphilic particles interacting with each other via hydrophobic forces in a solvent using continuum
modeling and numerical simulations.  Amphiphilic particles (such as lipid molecules) possess both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and bilayers of lipids) to shield their hydrophobic parts from contact with the solvent molecules (water).
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic force is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles for self-assembly in  soft matter \cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

\begin{wrapfigure}[16]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{The exterior domain $\Omega$ is the solvent region surrounding a
number of rigid particles $P_1,$ $P_2,$ $P_3,$ \dots 
The surface $S$ is the solvent-particle interface, indicated by the unit
normal $\mathbf{n}.$ }\label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the 
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM}.
This model addresses the major shortcomings of molecular dynamics (MD) and continuum approaches. Based on preliminary results (\S\ref{sec:preliminary_work})
the PIs propose to extend this HAP model to offer
 an alternative modeling methodology that leads to new mathematical ideas,
and is both physically accurate and computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles.
These results will facilitate optimal designs of smart materials of desirable properties by
tuning the geometry and properties of the amphiphilic particles.

\subsection{Hydrophobic Attraction Potential (HAP)}
\label{sec:HAP}

Based on the physical origin of hydrophobicity,  we have devised a functional, called the hydrophobic attraction potential (HAP), to
model hydrophobic forces \cite{Fu2018_SIAM}. The motivation for HAP concept stemmed from PI RR's 
earlier work on the problem of energy barriers in membrane fusion \cite{RyKlYaCo16,Chetal16}.
There, PI RR and collaborators resolved the long-standing issue of accounting for the energy of a fissure surface
during topological transitions, by applying a squared gradient mathematical theory 
for hydrophobic attraction between planar surfaces \cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977} to monolayer fissures.
Based on PI RR and YNY's coarse-grained membrane modeling work \cite{Fu2017}, 
the investigators initiated a collaboration to create a gradient theory for arbitrary collections 
of hydrophobic and amphiphilic particles. This new method 
eliminates the costly calculation of water by treating the solvent implicitly, 
and avoids complicated re-meshing schemes of continuum approaches
by utilizing a particle-based representation.

>>>>

\noindent
{\bf Collaborative Research: Mathematical modeling and simulations of
self-assembling amphiphilic particles in a solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
FSU), and Y.-N. Young (PI, NJIT)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use continuum modeling and
numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting with each other via hydrophobic forces
in a solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent molecules (water).
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic force is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles for self-assembly in  soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

\begin{wrapfigure}[16]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
  \todo[inline]{Change the $\mathbf{n}$ to a $\boldsymbol{\nu}$}
\caption{The exterior domain $\Omega$ is the solvent region surrounding a
number of rigid particles $P_1,$ $P_2,$ $P_3,$ \dots 
The surface $S$ is the solvent-particle interface, indicated by the unit
normal $\mathbf{n}.$ }\label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM} that is based
on the physical origin of hydrophobicity.  This model addresses the
major shortcomings of molecular dynamics (MD) and continuum approaches.
Based on preliminary results (\S\ref{sec:preliminary_work}) the PIs
propose to extend this HAP model to offer an alternative modeling
methodology that leads to new mathematical ideas, and is both physically
accurate and computationally practical.
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles.
These results will facilitate optimal designs of smart materials of desirable properties by
tuning the geometry and properties of the amphiphilic particles.

\subsection{Hydrophobic Attraction Potential (HAP)}
\label{sec:HAP}
%Based on the physical origin of hydrophobicity,  we have devised a
%functional, called the hydrophobic attraction potential (HAP), to model
%hydrophobic forces~\cite{Fu2018_SIAM}. 
The motivation for the HAP concept
stemmed from PI RR's earlier work on the problem of energy barriers in
membrane fusion~\cite{RyKlYaCo16,Chetal16}.  There, PI RR and
collaborators resolved the long-standing issue of accounting for the
energy of a fissure surface during topological transitions, by applying
a squared gradient mathematical theory for hydrophobic attraction
between planar surfaces~\cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977} to monolayer
fissures.  Based on PI RR and YNY's coarse-grained membrane modeling
work \cite{Fu2017}, the investigators initiated a collaboration to
create a gradient theory for arbitrary collections of hydrophobic and
amphiphilic particles. This new method eliminates the costly calculation
of water by treating the solvent implicitly, and avoids complicated
re-meshing schemes of continuum approaches by utilizing a particle-based
representation.
